---
title: "Project Uno"
author: "Anastasia Mikova"
date: "March 5, 2019"
output: html_document
---

# Structure of the Data

#### First, we are going to read the .csv file in R with `read.csv()` function.

```{r}
police_shootings <- read.csv(file = "shootings.csv", T, sep = ",")
# link, so you don't lose it!!! https://github.com/washingtonpost/data-police-shootings
```

#### Since we are going to use various packages, we need to install and load them.

```{r}
 #install.packages("dplyr")
 library(dplyr)
```

#### Then, we will look at the dimension and structure of the data with `dim()` and `str()` functions, respectively.

```{r}
dim(police_shootings)
```

```{r}
str(police_shootings)
```
###### The following dataset consists of 4054 observations of 14 variables 

#### Next, we are going to look at the head and the tail of the dataset.

```{r}
head(police_shootings)
```

```{r}
tail(police_shootings)
```

# Removing Missing Values

#### Next, we are going to remove all the missing values using `complete.cases()` function. We will call the new data frame 'police_shootings_full'. We are also going to remove the first column. 

```{r}
police_shootings_full <- police_shootings[complete.cases(police_shootings),]
police_shootings_full$id <- NULL
```

#### If we look at the dimension of the new data frame, we will notice that it consists of 3898 observations of 13 variables.

```{r}
dim(police_shootings_full)
```

```{r}
head(police_shootings_full, 3)
```

#### We are interested in how many suspects were unarmed when they were shot by the police. To find this out we can use `grep()` function for pattern matching.

```{r}
unarmed <- grep("unarmed", police_shootings_full$armed)
str(unarmed)
```
###### 260 out of 3898 suspects were unarmed when shot by the police.

#### We are going to do the same to find out how many suspects who were shot were black.

```{r}
black <- grep("B", police_shootings_full$race)
str(black)
```
###### 943 out of 3898 suspects were black.

#### `grepl()` function will give us a logical string of `TRUE` or `FALSE` values. Using `which()` function we can then select only `TRUE` values. This will be done to see how many suspects were white.

```{r}
white1 <- grepl("W", police_shootings_full$race)
white <- which(white1, T)
str(white)
```
###### 1891 out of 3898 suspects were white.